---
title: "hw9"
author: "Yasasvi"
date: "2024-04-25"
output: html_document
---

```{r}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(MASS)
library(ROCR)

data(iris)
```
###Model Development

Linear Discriminant Analysis (LDA) is performed on the dataset using the lda() function from the MASS package. LDA is a statistical method used for dimensionality reduction and classification. Here, the model is trained to predict the species of iris flowers based on their measurements.

###Prediction

The trained LDA model is used to make predictions on the same dataset (iris) using the predict() function. This step generates predictions for each observation in the dataset, including the predicted class and posterior probabilities.
```{r}
# LDA
lda_model <- lda(Species ~ ., data = iris)

lda_predictions <- predict(lda_model, newdata = iris)
lda_predictions

# predicted classes
predicted_classes <- lda_predictions$class
predicted_classes
```
###Model Evaluation

The accuracy of the model is calculated by comparing the predicted classes with the actual classes in the dataset. This gives an indication of how well the model performs in classifying iris flowers into their respective species.
```{r}
# Calculate accuracy
accuracy <- mean(predicted_classes == iris$Species)

# Summary of the LDA model and accuracy
cat("Summary of LDA Model:\n")
print(summary(lda_model))
print(lda_model)
cat("\nModel Accuracy:", round(accuracy * 100, 2), "%\n")
plot(lda_model)

# discriminant scores
lda_scores <- lda_predictions$x
lda_scores

# Convert the predicted probabilities to a data frame
predicted_probabilities <- as.data.frame(lda_predictions$posterior)
predicted_probabilities
```
###Model Accuracy

```{r}
pred <- prediction(predicted_probabilities[, "versicolor"], iris$Species == "versicolor")

roc.perf <- performance(pred, measure = "tpr", x.measure = "fpr")

auc.train <- performance(pred, measure = "auc")@y.values[[1]]

# ROC curve
plot(roc.perf, main = "ROC Curve", col = "blue", lwd = 2)
abline(a = 0, b = 1, lty = 2, col = "red")
text(x = 0.5, y = 0.3, paste("AUC = ", round(auc.train, 3), sep = ""))
```